### 2024-11-04 Manuscript/main_ecography_R1.tex
Update main_ecography_R1.tex

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.8\textwidth]{Figures/results_per_metric_per_stressor_4d.pdf}
\caption{\scriptsize{\textbf{Simulation results:} the y axes represent the different metrics tested (sorted by categories).
The different columns represent the different stressors. The x-axes represent the metric values centred on the random changes and scaled by the maximum value for each metric between the four stressors.
Negative and positive values signify that the metric score for the stressor of interest is lower/higher than the one from the random stressor.
The dots represent the median metric value, the full line their 50\% confidence interval (CI) and the dashed line their 95\% CI.
The colours are here to visually separate the metrics by categories (green = richness, yellow = divergence, blue = regularity); the colour gradient within each row corresponds to a removal of respectively 80\%, 60\%, 40\% and 20\% of the data (from top to bottom).
The grey line plots represent distributions of metric scores not clearly distinguishable from the random metric scores (paired t-test $p$ value $> 0.05$).
Grey lines in the background across the distributions of different removal amounts represent the fitted linear model ``centred and scaled metric score function of the amount of data removed''.
Dashed thin grey lines represent non-significant linear models (slope's $p$ value $> 0.05$).
The model results are available in supplementary table 3.
Similar figures are available in the supplementary materials for 2 and 8 dimensions.}}
\label{Fig:simulation_results}
\end{figure}
\bigskip

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.4\textwidth]{Figures/results_per_metric_per_stressor_emp.pdf}
\caption{\scriptsize{\textbf{Empirical results:} the metrics and the scaled changes are measured in the same way as in figure 2 (green = richness, yellow = divergence, blue = regularity).
Plain circles represent the metrics for the historic dataset (46\% removal) and fainted squares for the extant dataset (69\% removal).
\textbf{NOTE: for the \texttt{TPD} metrics, the metric is calculated using the 4 first dimensions only.}
}}
\label{Fig:empirical_results}
\end{figure}
\bigskip

@@

The empirical results are also intriguing, as it indicates that, at least for certain metrics, the loss of Hawaiian bird functional diversity is less than expected than if extinctions were random with respect to dimensions (Figure \ref{Fig:empirical_results}).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
This could partly be due to the proportion of passerine extinctions in Hawaii.
Although most global bird species are (or were, if accounting for anthropogenic extinctions) passerines (approximatively 58\%), the majority of global known anthropogenic extinctions (approx. 75\%) have been non-passerines, which tend to be larger and possess more unique morphological traits (thus resulting in relatively large reductions in functional diversity).
However, in Hawaii, where the same proportion of known species were passerines (approx. 58\%), the opposite is true: the majority of extinctions (approx. 60\%) have been of passerine species.
It is important to note that we are not implying that there existed little morphological variation within extinct Hawaiian passerines - indeed, many of these extinctions involved the honeycreepers, a famous adaptive radiation involving substantial evolution of beak morphology \citep{Walther2022} - but simply that this morphological variation is less than that observed across all birds.
But these results, given that the proportion of known Hawaiian species that are / were passerines roughly matches the proportion of extinct species that were passerines,  also indicate that these extinct species were more functionally similar than expected, thus relatively \textit{increasing} trait diversity (i.e. relative to random extinctions).
In addition, it is worth stressing that these analyses included marine species (which constitute less than 4\% of  known extinctions in Hawaii, while making up 11\% of the prehistoric assemblage) and different results may have been found if focusing exclusively on terrestrial birds.

@@

Another cautionary note could be made about our choice of scaling the metrics relative to a random trait space reduction (with the same percentage of reduction).
This choice allowed us to simulate a null model designed to test whether the mechanisms of interest were causing the observed change in pattern \citep{bausman2018modeling}.
For example, if removing 80\% of the edges of a distribution (equalising mechanism) was distinguishable or not from removing 80\% of a distribution randomly (null mechanism).
Some metrics (e.g. Functional evenness using the dendrogram method) were not able to distinguish between the null mechanism and the mechanism of interest (here the mechanism simulating competition).
In other words, our statistical question was ``Does metric X distinguish between removing N\% of data in a biassed way (the mechanism) and removing N\% of data randomly (the null)?''.
This definition of null hypothesis is also appropriate for the empirical data if the question is the same.
However, it is very likely that researchers will ask a more exciting or intriguing question based on these data.
For example, an interesting one could be ``Are the extinctions uniform across all birds in Hawaii?".
This legitimate question would thus require first a trait space (the pattern - what) and maybe some contrasting groups of interest like extinctions through time (the process - how) to answer the question (the mechanism - why).
Furthermore, the scaling of the metrics paired with the levels of removal (80\%, 60\%, 40\% and 20\%) creates an expected artifact similarity when comparing the random removals against the other ones: when less data is available overall, the metrics are more likely to be similar (e.g. if removing 99\% of the data, we expect the random and removal metric score to be nearly identical to the non-random ones).

>>>>

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.8\textwidth]{Figures/results_per_metric_per_stressor_4d.pdf}
\caption{\scriptsize{\textbf{Simulation results:} the y axes represent the different metrics tested (sorted by categories).
The different columns represent the different stressors. The x-axes represent the metric values centred on the random changes and scaled by the maximum value for each metric between the four stressors.
Negative and positive values signify that the metric score for the stressor of interest is lower/higher than the one from the random stressor.
The dots represent the median metric value, the full line their 50\% confidence interval (CI) and the dashed line their 95\% CI.
The colours are here to visually separate the metrics by categories (green = richness, yellow = divergence, blue = regularity); the colour gradient within each row corresponds to a removal of respectively 80\%, 60\%, 40\% and 20\% of the data (from top to bottom).
The grey line plots represent distributions of metric scores not clearly distinguishable from the random metric scores (paired t-test $p$ value $> 0.05$).
Grey lines in the background across the distributions of different removal amounts represent the fitted linear model ``centred and scaled metric score function of the amount of data removed''.
Dashed thin grey lines represent non-significant linear models (slope's $p$ value $> 0.05$).
The model results (including $R^{2}$ and t-statistics) are available in supplementary table 3.
Similar figures are available in the supplementary materials for 2 and 8 dimensions.}}
\label{Fig:simulation_results}
\end{figure}
\bigskip

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.4\textwidth]{Figures/results_per_metric_per_stressor_emp.pdf}
\caption{\scriptsize{\textbf{Empirical results:} the metrics and the scaled changes are measured in the same way as in figure 2 (green = richness, yellow = divergence, blue = regularity).
Plain circles represent the metrics for the species present on Hawaii before 1500 (historic data; 46\% removal) and fainted squares for the species currently present on Hawaii (extant data; 69\% removal).
\textbf{NOTE: for the \texttt{TPD} metrics, the metric is calculated using the 4 first dimensions only.}
The model results (including $R^{2}$ and t-statistics) are available in supplementary table 4.
}}
\label{Fig:empirical_results}
\end{figure}
\bigskip

@@

The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity since 1500 is not greater than expected by chance (at least for certain metrics; Figure \ref{Fig:empirical_results}) because one would expect that extinctions since 1500 target specific areas of functional diversity space (i.e. birds with specific trait combinations).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
% TODO:
% This illustrates the importance of setting the question + metric
% For example, for the divergence metrics using probability density, hypervolume or dendrogram methods, we recovered changes in trait space that are smaller than expected by chance for the species present before 1500 and larger than expected by change for the current species.
% This could lead to the interpretation that the trait space was not changed under any specific mechanism before 1500 (i.e. null mechanism) but that since then, specific mechanisms have been reducing specific areas of the trait space (e.g. filtering mechanism).
% However, when measuring divergence using the Rao's quadratic entropy based on distance matrices, we did not find that.
This could partly be due to the proportion of passerine extinctions in Hawaii.
Although most global bird species are (or were, if accounting for anthropogenic extinctions) passerines (approximatively 58\%), the majority of global known anthropogenic extinctions (approx. 75\%) have been non-passerines, which tend to be larger and possess more unique morphological traits (thus resulting in relatively large reductions in functional diversity).
However, in Hawaii, where the same proportion of known species were passerines (approx. 58\%), the opposite is true: the majority of extinctions (approx. 60\%) have been of passerine species.
It is important to note that we are not implying that there existed little morphological variation within extinct Hawaiian passerines - indeed, many of these extinctions involved the honeycreepers, a famous adaptive radiation involving substantial evolution of beak morphology \citep{Walther2022} - but simply that this morphological variation is less than that observed across all birds.
But these results, given that the proportion of known Hawaiian species that are / were passerines roughly matches the proportion of extinct species that were passerines,  also indicate that these extinct species were more functionally similar than expected, thus relatively \textit{increasing} trait diversity (i.e. relative to random extinctions).
In addition, it is worth stressing that these analyses included marine species (which constitute less than 4\% of  known extinctions in Hawaii, while making up 11\% of the prehistoric assemblage) and different results may have been found if focusing exclusively on terrestrial birds.

@@

Another cautionary note could be made about our choice of scaling the metrics relative to a random trait space reduction (with the same percentage of reduction).
This choice allowed us to simulate a null model designed to test whether the mechanisms of interest were causing the observed change in pattern \citep{bausman2018modeling}.
For example, if removing 80\% of the edges of a distribution was distinguishable or not from removing 80\% of a distribution randomly (e.g. facilitation mechanism measured as divergence using the probability density method - not distinguishable from null when removing 80\% or 60\% of the data; Figure \ref{Fig:simulations}).
Some metrics (e.g. Functional evenness using the dendrogram method) were not able to distinguish between the null mechanism and the mechanism of interest (here the mechanism simulating competition).
In other words, our statistical question was ``Does metric X distinguish between removing N\% of data in a biassed way (the mechanism) and removing N\% of data randomly (the null)?''.
This definition of null hypothesis is also appropriate for the empirical data if the question is the same.
However, it is very likely that researchers will ask a more exciting or intriguing question based on these data.
For example, an interesting one could be ``Are the extinctions uniform across all birds in Hawaii?".
This legitimate question would thus require first a trait space (the pattern - what) and maybe some contrasting groups of interest like extinctions through time (the process - how) to answer the question (the mechanism - why).
Furthermore, the scaling of the metrics paired with the levels of removal (80\%, 60\%, 40\% and 20\%) creates an expected artifact similarity when comparing the random removals against the other ones: when less data is available overall, the metrics are more likely to be similar (e.g. if removing 99\% of the data, we expect the random and removal metric score to be nearly identical to the non-random ones).
